Investigating magnetic dynamo evolution with \TESS\ field dwarfs

### 2017-10-06 ms.tex
addding references

Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star's rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star's surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\% or less) quasi-periodic changes in the stellar
brightness.
Based on results from \Kepler, using precise light curves available from the
\TESS\ mission, we expect to measure the rotation periods of around 20\% of FG
stars and 80\% of KM stars in the \tess\ Full Frame Images (FFIs)\footnote{The
difference between these percentages comes from a dramatic increase in
magnetic activity, and therefore more easily measured rotation periods,
towards lower masses.}.
For many of these stars it will be possible to infer an age from their
rotation periods via gyrochronology.
Some outstanding mysteries regarding the nature of magnetic braking were
revealed, but thus-far unanswered, by \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars
\citep{van-saders2016}.
To answer these questions {\bf we propose to provide a catalog of rotation
periods for \TESS\ Full Frame Image (FFI) stars.}

@@

\citet{oh2016} identified over 10,000 comoving pairs of stars using proper
motions from the first \Gaia\ data release.
Of these, 34 dwarfs and subgiants in 17 pairs have \RAVE\ radial velocities
consistent with being comoving and are already in the Candidate Target List.
Since these stars would make interesting targets for both asteroseismology and
exoplanet search, we include a list with this proposal.
The stars in these pairs are likely recently disrupted binaries so we
expect them to have similar ages and measuring their rotation periods will
therefore provide a test of gyrochronology.
An additional 5000 bright comoving pair candidates lie in the southern
ecliptic hemisphere and will be ideal FFI targets for gyrochronology studies.
{\bf We will use the rotation periods of comoving stars to quantify the
accuracy and precision of gyrochronology.}

@@

Based on \citet{sullivan2015}, we expect around 10 million \TESS\ FFI
targets brighter than 15th magnitude in the southern ecliptic hemisphere.
We intend to prioritize certain groups of stars for period analysis,
calculating a simple (and computationally fast) Lomb-Scargle periodogram for
the low priority objects and running a full probabilistic, Gaussian process
period analysis on the high priority targets.
High priority targets will include: comoving pairs, stars in the CVZ
(especially those with spectra), stars in TGAS, stars at large distances, plus
any potential asteroseismic dwarfs and exoplanet hosts.
The number of high priority targets with probabilistic rotation periods will
be limited by computation power, but we conservatively estimate that we can
produce full rotation period posterior PDFs for $\sim$ 30,000 stars.
Based on the expected photometric precision of \tess\ \citep{sullivan2015}, we
expect to be able to measure rotation periods for FFI stars down to 15th
magnitude.
We note, however that not all stars will have measurable rotation periods,
even if they are bright enough not all stars show rotational variability in
their light curves either because they are pole-on or inactive.
We will vet incorrect rotation period measurements using a combination of a
Lomb-Scargle periodogram and an autocorrelation function by
training on a set of light curves with a range of variability amplitudes and
periods.
However, based on the high success rates of a Lomb-Scargle periodogram method
with high-pass filtered light curve and the Gaussian process method,
demonstrated in \citet{angus2017}, we expect a low rate of incorrect rotation
periods.

>>>>

Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star's rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star's surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\% or less) quasi-periodic changes in the stellar
brightness.
Based on results from \Kepler, using precise light curves available from the
\TESS\ mission, we expect to measure the rotation periods of around 20\% of FG
stars and 80\% of KM stars in the \tess\ Full Frame Images (FFIs)\footnote{The
difference between these percentages comes from a dramatic increase in
magnetic activity, and therefore more easily measured rotation periods,
towards lower masses.}.
For many of these stars it will be possible to infer an age from their
rotation periods via gyrochronology.
Some outstanding mysteries regarding the nature of magnetic braking were
revealed, but thus-far unanswered, by \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars
\citep{van-saders2016}.
To answer these questions {\bf we propose to provide a catalog of $\sim$ 1
million rotation periods for \TESS\ Full Frame Image (FFI) stars.}

@@

\citet{oh2016} identified over 10,000 comoving pairs of stars using proper
motions from the first \Gaia\ data release.
Of these, 18 dwarfs and subgiants in 9 pairs have \RAVE\ radial velocities
consistent with being comoving and are already in the Candidate Target List.
Following \citet{campante2017}, these stars have oscillation detection
probabilities greater than 50\%, based on their \RAVE-on T$_{\mathrm{eff}}$,
$\log g$s and radii \citep{casey2017}.
Since these stars would make interesting targets for both asteroseismology and
exoplanet search, we include a list with this proposal.
The stars in these pairs are likely recently disrupted binaries so we
expect them to have similar ages and measuring their rotation periods will
therefore provide a test of gyrochronology.
An additional 5000 bright comoving pair candidates lie in the southern
ecliptic hemisphere and will be ideal FFI targets for gyrochronology studies.
{\bf We will use the rotation periods of comoving stars to quantify the
accuracy and precision of gyrochronology.}

@@

Based on \citet{sullivan2015}, we expect around 10 million \TESS\ FFI
targets brighter than 15th magnitude in the southern ecliptic hemisphere.
We intend to prioritize certain groups of stars for period analysis,
calculating a simple (and computationally fast) Lomb-Scargle periodogram for
the low priority objects and running a full probabilistic, Gaussian process
period analysis on the high priority targets.
High priority targets will include: comoving pairs, stars in the CVZ
(especially those with spectra), stars in TGAS, stars at large distances, plus
any potential asteroseismic dwarfs and exoplanet hosts.
The number of high priority targets with probabilistic rotation periods will
be limited by computation power, but we conservatively estimate that we can
produce full rotation period posterior PDFs for $\sim$ 30,000 stars.
Based on the expected photometric precision of \tess\ \citep{sullivan2015}, we
expect to be able to measure rotation periods for FFI stars down to 15th
magnitude.
We note, however that not all stars will have measurable rotation periods,
even if they are bright enough not all stars show rotational variability in
their light curves either because they are pole-on or inactive.
We will vet incorrect rotation period measurements using a combination of a
Lomb-Scargle periodogram and an autocorrelation function by
training on a set of light curves with a range of variability amplitudes and
periods.
Based on the fraction of stars with measurable rotation periods in
\citet{mcquillan2014}, and the more restrictive baseline of \tess\ relative to
\kepler, we predict that the final catalog will contain around one million
rotation periods.
